Better fusion for filters

### 2014-04-01 2014betterfusionforfilters/section/04-ILP.tex
ilp intro

\section{Integer linear program formulation}

>>>>

\section{Integer linear program formulation}
While the different combinators are all implemented differently, certain aspects are shared between all; they operate on input arrays, use scalar variables in their worker functions, and produce output.
To simplify creation of the integer linear program formulation, we first convert a function to a dependency graph.
The dependency graph has a node for each combinator, and edges between two combinators when one uses another's output.
After the graph is created, it is converted to an integer linear program with an objective of the least memory traffic and loops, solved, and the solution converted to a clustering.

When creating the graph, it is important to note that not all loops can be fused together; firstly, loops of different size cannot be fused.
Secondly, two loops cannot be fused together if an iteration in one loop relies on the output of a later iteration in another loop.
For example, a @map@ may not be fused with a @fold@ if the @map@'s worker function uses the result of the @fold@.
This fusion restriction is encoded as a \emph{fusion-preventing} edge between two combinators.
Other edges are \emph{fusible}, and may be fused together.
If they are not fused together for whatever reason, the \emph{from} combinator must be scheduled before the \emph{to} combinator.

The dependency graph is translated to an integer linear program.
The integer linear program has some integral, boolean and real variables, an objective function to minimise, and a set of constraints that the variables must conform to.
Finding a variable assignment that satisfies the constraints and is the minimal objective function is NP-hard, but existing ILP solvers tend to be sufficient for realistic problem sizes.
For larger problems, we can find an approximate answer, within say $10\%$ of the optimal answer, which still gives the exact answer for small problems.
